\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-21 paper/main.tex
.

% \subsection{Baseline model}
Furthermore, we hypothesize that some of the predictions made by our
model are the result of the biases in drug--target interaction
datasets where hub proteins may have a large number of interactions,
or certain drugs interact with many proteins, and these may be easier
to predict than others. We design a ``na\"ive'' baseline model that
predicts the same list of proteins for each drug based only on the
number of known drug--target interactions for a protein. Formally,
given lists $D$ and $P$ of drugs and proteins and a set of known
interactions $\mathcal{I} := \{(d_i, p_i) \}$, we construct an
interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein pairs whether there is a known
interaction or not. We now rank all proteins $p_j\in P$ descending by
their number of drug interactors by summing over the columns of
$M_{ij}$ and ranking these sums:
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
Our ``na\"ive'' predictor $P_k$ predicts all drugs to interact with
the top $k$ targets with respect to the introduced ranking:
\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in Top_{k}(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$.

>>>>

% \subsection{Baseline model}
Furthermore, we hypothesize that it may be possible to exploit biases
in drug--target interactions to achieve relatively high prediction
performance without obtaining a biologically meaningful signal.
For example, hub proteins may have a large number of interactions,
or certain drugs interact with many proteins, and preferentially
predicting these interactions may increase predictive performance even
in the absence of any biological features.
To test this hypothesis, we design a ``na\"ive'' baseline model that
predicts the same list of proteins for each drug based only on the
number of known drug--target interactions for a protein. Formally,
given lists $D$ and $P$ of drugs and proteins and a set of known
interactions $\mathcal{I} := \{(d_i, p_i) \}$, we construct an
interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein pairs whether there is a known
interaction or not. We now rank all proteins $p_j\in P$ descending by
their number of drug interactors by summing over the columns of
$M_{ij}$ and ranking these sums:
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
Our ``na\"ive'' predictor $P_k$ predicts all drugs to interact with
the top $k$ targets with respect to the introduced ranking:
\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in Top_{k}(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$.
